Stateless Model Checking with Data-Race Preemption Points

### 2016-07-20 oopsla-16/overview.tex
garth revisions

@@

	{\bf Variants of Happens-Before.}
	Most prior work focuses on {\em Happens-Before} (HB) \cite{lamport-clocks,djit,fasttrack} as the order relation between accesses.
\cite{predictive-dr} and \cite{hybriddatarace} identify a problem with this approach:
it cannot identify access pairs separated by an unrelated lock operation which could race in an alternate interleaving,
as shown in the example program in Figure~\ref{fig:hb-example}(a).
We call such unreported access pairs {\em false negatives}.
However, consider the similar program in Figure~\ref{fig:hb-example}(b),
in which the access pair ceases to exist in the alternate interleaving.
\cite{hybriddatarace} introduces the {\em Limited HB} relation,
which will report such potential races
by considering only blocking operations like {\tt cond\_wait} to enforce the order.
Limited HB will report all potential races, avoiding \revisionx{many} false negatives \cite{tsan},
but at the cost of necessarily reporting some such {\em false positives}.

@@

\revisionx{
{\bf Philosophy of bugs.}
% two camps: all races are bugs; citations:
% and, races must be classified; citations: replay analysis, portend,
While there is a vast body of work on how to detect data races to begin with,
judging data races once found is a matter of philosophical debate unto itself.
Some recent tools classify races depending on how they impact program behaviour \cite{recordreplaydrs,portend}, overlooking {\em benign} races in search of more dangerous ones.
%In particular,
\cite{racerx} acknowledges that a program may have too many races for a user to worry about,
so bug reports must be prioritized by severity of effects. %how likely they may cause failures or security vulnerabilities.
%
However, other prior work argues that data races are always bugs \cite{miscompile-benign,data-races-are-evil},
largely due to the possibility of compiler or hardware reordering of racy accesses.
%
C++ declares any race between two non-atomic locations to be undefined behaviour \cite{cpp-foundation}.
We take the former camp:
we consider a data race a bug only when it results in a visible failure state (e.g., crash or deadlock).
(From a C++ perspective, we assume all concurrent accesses are implemented by {\tt std::atomic}.)
We bypass concerns of compiler reordering by checking programs at the executable level;
for a study of hardware reordering in the context of DPOR, we refer the reader to \cite{tsopso}.
}

>>>>

\revisionx{
{\bf Recent advances.}
Like many prior checkers \cite{chess,dbug-ssv,inspect,portend}, ours implements DPOR for sequentially-consistent hardware.
In the worst case, these tools may all suffer false negatives as they miss weak-memory-only bugs.
Recently, Zhang et al. \cite{tsopso} introduced a new formalism with which DPOR can control weak memory nondeterminism,
such as reordering store buffers.
Iterative Deepening could use this new technique,
%We expect this technique would apply directly with Iterative Deepening,
provided a model checker which implements such reorderings (not yet available to us).
%We believe this assumption is reasonable for the scope of this paper because the main property we require of DPOR is that
%A recent paper extended DPOR to support weaker memory models \cite{tsopso}, so our proofs could likewise be extended with that version of DPOR in future work.

Maximal Causality Reduction (MCR) \cite{mcr}, a reduction algorithm which may replace DPOR,
has also recently shown promising performance improvements over prior model checkers.
We expect Iterative Deepening to be compatible with MCR,
and look forward to evaluating the combination when an MCR implementation becomes widely available.
}

@@

	{\bf Variants of Happens-Before.}
	Most prior work focuses on {\em Happens-Before} (HB) \cite{lamport-clocks,djit,fasttrack} as the order relation between accesses.
\cite{predictive-dr} and \cite{hybriddatarace} identify a problem with this approach:
it cannot identify access pairs separated by an unrelated lock operation which could race in an alternate interleaving.
%as shown in the example program in Figure~\ref{fig:hb-example}(a).
\revisionx{Figure~\ref{fig:hb-example}(a) shows a contrived example program in which HB masks the potential race.}
We call such unreported access pairs {\em false negatives}.
However, consider the similar program in Figure~\ref{fig:hb-example}(b),
in which the access pair ceases to exist in the alternate interleaving.
O'Callahan et al. \cite{hybriddatarace} introduced the {\em Limited HB} relation,
which will report such potential races
by considering only blocking operations like {\tt cond\_wait} to enforce the order.
Limited HB will report all potential races, avoiding \revisionx{many} false negatives \cite{tsan},
but at the cost of necessarily reporting some such {\em false positives}.

@@

\revisionx{
{\bf Philosophy of bugs.}
% two camps: all races are bugs; citations:
% and, races must be classified; citations: replay analysis, portend,
While there is a vast body of work on how to detect data races to begin with,
judging data races once found is a matter of philosophical debate unto itself.
Some recent tools classify races depending on how they impact program behaviour \cite{recordreplaydrs,portend}, overlooking {\em benign} races in search of more dangerous ones.
%In particular,
\cite{racerx} acknowledges that a program may have too many races for a user to worry about,
so bug reports must be prioritized by severity of effects. %how likely they may cause failures or security vulnerabilities.
%
However, other prior work argues that data races are always bugs \cite{miscompile-benign,data-races-are-evil},
largely due to the possibility of compiler or hardware reordering of racy accesses.
%
(C++ declares any race between two non-atomic locations to be undefined behaviour \cite{cpp-foundation}.)
We take the former camp:
we consider a data race a bug only when it results in a visible failure state (e.g., crash or deadlock).
(From a C++ perspective, we assume all concurrent accesses are implemented by {\tt std::atomic}.)
We bypass concerns of compiler reordering by checking programs at the executable level;
for a study of hardware reordering in the context of DPOR, we refer the reader to \cite{tsopso}.
}

### 2016-07-20 oopsla-16/soundness.tex
garth revisions

\revisionx{These proofs are built on a DPOR algorithm definition which assumes sequentially-consistent memory hardware.
%
% TODO: Aaargh, can't think of how to actually justify this.
%We believe this assumption is reasonable for the scope of this paper because the main property we require of DPOR is that 
A recent paper extended DPOR to support weaker memory models \cite{tsopso}, so our proofs could likewise be extended with that version of DPOR in future work.
We also assume the Limited HB definition for the data-race analysis, leaving the case for precise HB to future work.
}

>>>>

\revisionx{These proofs are built on a DPOR algorithm definition which assumes sequentially-consistent memory hardware,
as discussed in \sect{\ref{sec:overview-sssmc}}.
We also assume the Limited HB definition
for the data-race analysis, as discussed in
\sect{\ref{sec:overview-dr}}.
%leaving the case for precise HB to future work.
}
